---
title: "Machine Learning Final"
author: "Melissa Miller"
date: "February 18, 2018"
output: html_document
---

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

The first part of this is to load the data into the program.  

```{r}

train_data <- read.csv("C:/Users/mloun/Documents/pml-training.csv")
test_data <- read.csv("C:/Users/mloun/Documents/pml-testing.csv")
dim(train_data)
```

**Cleaning the DAta** 
The data needs to be cleaned there are non numeric values and there are NA's through out the data.  


```{r}
na_factor<-apply(train_data,2,function(col)sum(is.na(col))/length(col))*100
na_factor=as.data.frame(na_factor)
na_factor$colname=rownames(na_factor)
rownames(na_factor)=NULL
#plot_na=ggplot(subset(na_factor,na_factor>95), aes(colname, na_factor))+geom_bar(stat="identity")+coord_flip()
#plot_na
subset_na_95=subset(na_factor, na_factor>95)
na_thrash <- !names(train_data) %in% subset_na_95$colname
train_data_trans <- train_data[,na_thrash] 
```

```{r}
still<-c("kurtosis_roll_belt",  "kurtosis_picth_belt",  "kurtosis_yaw_belt",    "skewness_roll_belt",   "skewness_roll_belt.1", "skewness_yaw_belt", "max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "kurtosis_roll_arm", "kurtosis_picth_arm",   "kurtosis_yaw_arm", "skewness_roll_arm",    "skewness_pitch_arm",   "skewness_yaw_arm", "kurtosis_roll_arm",    "kurtosis_picth_arm",   "kurtosis_yaw_arm", "skewness_roll_arm",    "skewness_pitch_arm",   "skewness_yaw_arm", "kurtosis_roll_dumbbell",   "kurtosis_picth_dumbbell",  "kurtosis_yaw_dumbbell",    "skewness_roll_dumbbell",   "skewness_pitch_dumbbell",  "skewness_yaw_dumbbell" , "max_yaw_dumbbell" ,"min_yaw_dumbbell" ,"amplitude_yaw_dumbbell", "max_yaw_forearm", "max_yaw_forearm", "min_yaw_forearm","kurtosis_picth_forearm", "skewness_yaw_forearm","amplitude_yaw_forearm","kurtosis_picth_forearm","kurtosis_yaw_forearm","skewness_roll_forearm","kurtosis_roll_forearm","skewness_pitch_forearm")
still_thrash <- !names(train_data_trans) %in% still
train_data_trans_1 <- train_data_trans[,still_thrash] 
train_data_trans_1<- train_data_trans_1[,-c(1:7)]
train_data_trans_1$classe=as.factor(train_data$classe)
```

**Cross Validation and splitting the data** 
Cross validation allows us to split the large data into different groups in order to run analysis.  
```{r}
set.seed(77777)
library(caret)

training <- createDataPartition(y=train_data_trans_1$classe,
                               p=0.75,list=FALSE)
training_data <- train_data_trans_1[training,]
validation_data <- train_data_trans_1[-training,]
```
**Coorelation**
This allows us to look at varaibles and determine if a relationship exists. Specifically the graph looks like there is a positive coorelation between variables.   
```{r}
m <- abs(cor(training_data[,-53]))
diag(m) <-0
which(m>0.8, arr.ind=T)


plot(training_data[,21], training_data[,24])
```

```{r}
preProc<- preProcess (training_data[,-53], method="pca", thresh=0.95)
training_data_preprocess<- predict (preProc, training_data[,-53])
training_data_preprocess$classe=training_data$classe

validation_data_preprocess <- predict (preProc, validation_data[,-53])
validation_data_preprocess$classe=validation_data$classe
```
By validating the data we are able to begin our model building.  

**Model building**
 The random Forest algorithm tests the sensativity between trees.  It is a great tool for prediction.  This will allow us to see how sensative the variables are to one another.  
 
```{r}
 library(randomForest)
fit<-randomForest(classe~., data=training_data)
print(fit)

confusionMatrix(table(predict(fit, training_data), training_data$classe))

confusionMatrix(table(predict(fit, validation_data), validation_data$classe))
```

Looking at these tables, we can see that there is a 95.5% accuracy rating! 

```{r}
fit_preprocess<-randomForest(classe~., data=training_data_preprocess)
print(fit_preprocess)

confusionMatrix(table(predict(fit_preprocess, training_data_preprocess), training_data_preprocess$classe))

confusionMatrix(table(predict(fit_preprocess, validation_data_preprocess), validation_data_preprocess$classe))
```
Looking at the tables we can see a high level of sensitivity (about 97%) and a relatively small confidence interval.  

**Prediction** 
```{r}
test_data <- read.csv("C:/Users/mloun/Documents/pml-testing.csv")
train_lables<- colnames(test_data) %in% colnames(training_data)
test_data_filtered<- test_data[train_lables]
setdiff(names(training_data), names(test_data_filtered))

pred_test<-predict(fit, test_data_filtered)
pred_test
```

